---
title: "HOMEWORK2"
author: "Lucy Lockwood"
date: "October 1, 2015"
output: html_document
---

### Lockwood Homework #2
~~~
Thank you for finally helping me to understand the relationship between R,
R scriptfiles, R Markdown, and Knir.
~~~

```{r}
# set up working directory
#setwd("~/R/R_Projects/BIOL_653")

# Set up needed libraries
library(dplyr)
library(ggplot2)
library(tidyr)

# set up data set being used
library(gapminder)
```

### PART I - Make these plots!


#### Plot 1 - use points and color to identify continents

```{r}
ggplot(data= gapminder, aes(x= gdpPercap, y= lifeExp, color= continent)) + 
  geom_point(aes(shape= continent))
```
 

#### Plot 2 - Since the per capita gpd has a very large range,
####  use a transformation to linearise the data.
```{r}
ggplot(data= gapminder, aes(x= log(gdpPercap), y= lifeExp, color= continent)) + 
    geom_point(aes(shape= continent)) +
    xlab("gdpPercap")
```

#### Plot 3 - include a simple linear fit to the transformed data
```{r}
ggplot(data= gapminder, aes(x= log(gdpPercap), y= lifeExp, color= continent)) + 
  geom_point(aes(shape= continent)) +
  xlab("gdpPercap") +
  # geom_smooth(color = "blue", method= "lm", se= FALSE)
  geom_smooth(color = "blue", method= "lm")
```


#### Plot 4 - plot the density functions of life expectancy for each continent
```{r}
ggplot(data= gapminder, aes(x= lifeExp,y= , color= continent, fill= continent)) + 
 # geom_density(aes(y= log(gdpPercap))) +
  geom_density(color = "black", alpha = 0.6) + 
  facet_wrap(~year) 
```

#### Plot 5 - for a bonus, fix the bottom continent labels
```{r}
ggplot(data= gapminder, aes(x= continent,y= lifeExp, color= continent)) + 
  geom_boxplot() + 
  facet_wrap(~year) 

ggplot(data= gapminder, aes(x= continent,y= lifeExp, color= continent)) + 
  geom_boxplot() + 
  facet_wrap(~year) +
  theme(axis.text.x = element_text(angle = 50, size=10, vjust = 0.5))
```

### PART TWO - Make these plots and then some!

#### use dplyr to calculate summary statistics
#### the function subset() may be useful

#### Plot 6
```{r}
ggplot(data= gapminder, aes(x= lifeExp,y= , color= continent, fill= continent)) + 
  # geom_density(aes(y= log(gdpPercap))) +
  geom_density(color = "black", alpha = 0.6)
```


#### Plot 7 - Plot 7 - plot the mean life expectancy on a density plot for Asia
```{r}
asia_data <- filter(gapminder, continent == "Asia")

ggplot(data= asia_data,
       aes(x = lifeExp,y = , color = continent, fill = continent)) + 
  geom_density(color = "black", alpha = 0.6, fill = "green") +
  geom_vline(xintercept = mean(asia_data$lifeExp)) +
  ggtitle("Life Expectancy in Asia")
```
  

#### 8) Plot 8
 
#### a. Create a data.frame of the mean life expectancies for each continent.
####  You should end up with a data.frame like this
```{r}
meanLife <- gapminder %>%
  group_by(continent) %>%
  summarise(avg_life= mean(lifeExp))
meanLife
```

#### b. Plot the density plot of life expectancies for each continent and
####  draw a vertical line to mark the mean life expectancy for each continent.

```{r}    
meanLife <- gapminder %>%
  group_by(continent) %>%
  summarise(avg_life= mean(lifeExp))
  
ggplot(data = gapminder, aes(x=lifeExp, color = continent, fill = continent)) +
  geom_density(color = "black", alpha = 0.6) +
  facet_wrap(~continent) +
  geom_vline(data=meanLife, aes(xintercept = avg_life))
```   

### Part III - Troubleshooting practice

```{r}
hw_gapminder <- read.csv('hw_gapminder.csv')
```
####The problem here is:
  a typo of an added "e" on field name lifeExp
```{r}
# mean_lifeExp <- mean(hw_gapminder$lifeExpe)
# Corrected code line:
mean_lifeExp <- mean(hw_gapminder$lifeExp)
mean_lifeExp
```

####The problem here is:
  missing "c" before second string providing the "y" values for the list array
```{r}
#small_set <- hw_gapminder[c(1, 2, 3, 4, 1300:1304), +
#  ('country', 'continent', 'year')]
small_set <- hw_gapminder[c(1, 2, 3, 4, 1300:1304), c('country', 'continent', 'year')]
small_set
```
####The problem here is:
  NA values in the dataset but with the default na.rm = FALSE the NAs are not
  stripped out and mess up the mean calculation.
```{r}  
mean_gdp <- mean(hw_gapminder$gdpPercap)
# Corrected code: Changed argument na.rm = TRUE
mean_gdp <- mean(hw_gapminder$gdpPercap, na.rm = TRUE)
mean_gdp
```

####The problem here is:
  missing second "=" sign to make equivalent to rather than to assign the value
  
```{r}
#max_country <- hw_gapminder$country +
# [which(hw_gapminder$lifeExp = max(hw_gapminder$lifeExp))]
# Corrected code:
max_country <- hw_gapminder$country[which(hw_gapminder$lifeExp == max(hw_gapminder$lifeExp))]
max_country
```





